---
title: "ESM 244 Lab 1"
author: "Hanna Buechi"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

- basic review (data wrangling)
- intro to spatial data in R with sf

1. Attach packages

```{r load_packages}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

# Issue: installed sf() and tmap() but could not load

```

2. Review on reading, exploring, wrangling, and visualizing data

a. Read in the data
```{r}

wine <- read_csv("wine_data.csv") # this is already in my working directory (to the right) that I cloned from GitHub

# View(wine) my tmap is taking a long time to install, so I can't run code at the moment. I'm keeping track of what Allison does until I can run my own code.
# I get an error that I'm not in the right working directory ('/Users/hannabuechi/Desktop' --> should be in my Github one) -- how do I make sure I set the working directory from the beginning

```


b. Wrangling
```{r}

wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% # makes a new column, changing the name to state
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% # filters out rows that match a condition
  mutate(ppd = points/price) %>% # new column named ppd
  arrange(-ppd)

```

c. Use group_by() and summarize() to find summary statistics by group
```{r}

wine_summary <- wine_new %>% # make a new dataframe (tibble??)
  group_by(state) %>% # create groups in R's brain that we wont' see but do exist to calculate things based on these level
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), # remove NA and calculate mean, WA has highest points per price
    max_ppd = max(ppd, na.rm = TRUE),
    min_ppd = min(ppd, na.rm = TRUE),
    med_ppd = median(ppd, na.rm = TRUE)
  )

```

d. A few gg-graphics

```{r}

wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap(~state) # there are some catches with what this is representing by state... explore

# Bare min for ggplot: ggplot(), data, aes() for x/y, type of graph

wine_hists
```

New graph: scatterplot graph of price vs. point, with color determined by state, adding a linear model trendline on top of it

```{r}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) + # points use "color", bars use "fill"
  geom_smooth(color = "black", method = "lm", se = FALSE) + # puts trendlines in by OLS (ordinary least sq regression)
  scale_color_manual(values = c("royalblue","mediumorchid1","orange")) +
  facet_wrap(~state) + # so that WA shows up
  theme_light()

```

2. California counties (spatial data) using sf

```{r}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016") # st for spatiotemporal; "." if we are already in our working directory # prefix for all of the associated spatial data files, ie. .cpg, .dbf, .shp, etc

# plot(ca_counties) # don't do this, it might take a long time! It's trying to map and colorcode

# View(ca_counties) to see attributes

# sf uses sticky geometries ; KEEPS GEOMETRIES OF DATAFRAME --> treat it like a normal dataframe # significant to "keep geometry" --> retains all of the polygon information

ca_land <- ca_counties %>% 
  select(NAME, ALAND) # simplified information that retains polygon information

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

ca_df <- full_join(ca_land, ca_pop_inc) %>%  # keeps every row even if it doesn't have a match in the other dataframe
  select(NAME, MedFamilyIncome)

```

d. Use geom_sf to make a map in ggplot

```{r}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% # needs a particular coordinate system
  addTiles() %>% 
  addPolygons() # interactive map of polygons

```










